Edge effects and optimal field sizes in Alberta

### 2021-06-15 Manuscript/manuscript.Rmd
Worked on Intro and Methods

- Preserving SNL around fields is important for agriculture and conservation
  - Ecosystem services
  - Act as reservoirs for beneficial insects
  - Microclimate zones
  - Poorly studied in North America
  - Studies tend to be limited in scope (e.g. only certain organisms/crops) and applicability

- Ecosystem services can influence both the mean and variability of yield in agroecosystems
  - Typically only averages (means) are considered, but higher stability (lower variance) in yield is also valuable

- Field size and field boundaries are directly related to the conservation of SNL in agroecosystems
  - Size examples: wheat study from the UK
  - Boundary examples: flower strip studies, hedgerows
  - In North America, there are financial incentives to make fields large and homogeneous, especially with large agricultural equipment
  - Crop edge effects cause low yields at the margins of fields because of late emergence, poor microclimate, and competition with weeds
  - At the same time, ecosystem services should decay with distance from edge, so crops at the centre of a large field will not benefit from ecosystem services
  - This suggests that yield may be maximized at distances between the edge and centre of the field, where ecosystem services cancel out negative edge effects

  <!-- - This suggests a ``goldilocks" field size, where negative edge effects are canceled out by ecosystem services -->

- Precision yield data holds enormous promise for agronomy
  - Limited because of:
  - Lack of standardized formats
  - Yearly calibration data
  - Clear statistical protocols

@@

  - where:
  1. Polygon Size = distance traveled x width of header bar ($m^2$)
  2. Distance from Edge = distance from field edge type _i_ (m)
  3. Easting, Northing = distance from centroid of field (m)
  4. Harvest Sequence = order of harvest within field (1--N points)
  5. f(x,b) = penalized thin-plate regression spline, where x is the predictor and b is the number of basis dimensions

- This model accounts for a) combine speed differences, b) edge effects, c) within-field spatial variation not related to edges, and d) temporal differences in harvest 
  - Spatial or temporal variation is typically modelled using a Gaussian Process Model (e.g. Kriging), but this was computationally infeasible with 50,000 data points per field
  - Penalized regression splines offer a compromise, as they account for nonlinear ``wiggly" relationships in the same way as Gaussian processes but with substantially reduced computation time
  - The number of basis dimensions was checked with the _gam.check_ function from _mgcv_
  - A smoother was not used for polygon size because the relationship was easily approximated with a log-linear relationship using a single slope term
  - All models were fit in _R_ using the _mgcv_ library, and figures were created with _ggplot2_ and _ggpubr_

- Most models still had a large amount of spatio-temporal autocorrelation in residuals
  - We basically just ignored this

- To combine results from all field-level models while also accounting for field-level uncertainty, we used an approach similar to bootstrapping of hierarchical mixed effects models
  - The approach was as follows
  1. Extract posterior samples of smoother parameters from each field-level model
  2. Fit meta-model of all smoothers, and save parameters
  3. Repeat 1000 times, and calculate credible intervals (10-90th percentiles of draws)

- We compared this

>>>>

- Intensive agricultural production has increased over the last 100 years, and agricultural land now makes up over a third of ice-free land on Earth
  - This has allows increases in human population and increased (global) stability in production
  - However, this is not without cost, as higher-diversity non-crops are converted to lower-diversity crops, resulting in loss of habitat and overall biodiversity of non-target organisms
  - Maintaining both biodiversity and production in agroecosystems represents a seldom-considered goal of conservationists and agronomists, and hold the potential for win-win scenarios
  - Key to this is the preservation of semi-natural land (SNL), which represents the interface between crops and non-crops within agroecosystems

- SNL in and around crops is important for both agricultural production and conservation
  - They are habitat for mobile organisms, and can therefore act as sources of ecosystem services such as pollination or pest control
  - They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed
  - Unfortunately, most of the research is concentrated in Europe, and they tend to be less-studied on other continents
  - In particular, North American agroecosystems have larger fields, and different varieties and agronomic practices, all of which could negate effects of SNL

<!-- - SNL has conflicting effects on crop yield because of ecosystem service provision as well as negative edge effects -->
- SNL may affect yields at intermediate distances, depending on the spatial scale at which ecosystem services operate
  - Edge effects cause low yields at the edge of crops because of sparse or late seedling emergence, poor microclimate, and competition with weeds
  - At the same time, the centre of large fields will not receive ecosystem services if they decay with distance from edges
  - For example, pollination services from central place foragers that nest in SNL but forage in crops drops rapidly with distance
  - Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects 
  - This suggests a ``goldilocks" field size, where negative edge effects are canceled out by ecosystem services

<!-- - Less-considered are ecosystem ``disservices", where weeds and pests  -->

<!-- - In North America, there are financial incentives to make fields large, homogeneous, and with simple geometry -->

- Studies of SNL effects on crop yield also suffer from limited scope (e.g. few crop types) and small sample sizes, limiting inference and reducing generality (but see wheat study)
  - For this reason, large-scale precision yield data holds enormous promise for agronomy, as it allows 
  - However, its use is limited for several reasons: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) unfamiliarity with spatial statistics
  - Ecosystem services can influence both the mean and variability of yield in agroecosystems
  - Typically only averages (means) are considered, but higher stability (lower variance) in yield can also be valuable
  - Size examples: wheat study from the UK
  - There are few studies of yield variability (only those with large datasets), but precision yield data opens up the possibility of modeling within-field variability, as well as average yield

@@

  - where:
  1. Polygon Size = distance traveled $\times$ width of header bar ($m^2$)
  2. Distance from Edge = distance from field edge type _i_ (m)
  3. Easting, Northing = distance from centroid of field (m)
  4. Harvest Sequence = order of harvest within field (1--N points)
  5. f(x,b) = penalized thin-plate regression spline, where x is the predictor and b is the number of basis dimensions

- In addition to modeling edge effects (our variable of interest), this model also accounts for a) differences in combine speed, b) within-field spatial variation not related to edges, and d) shifts in combine accuracy during harvest
  - Spatial or temporal variation is typically modelled using a Gaussian Process Model (e.g. Kriging) or Stochastic Partial Differential Equations (e.g. INLA), but this was computationally infeasible with 50,000 data points per field
  - Penalized splines offer a compromise, as they account for nonlinear ``wiggly" relationships in the same way as Gaussian processes but with substantially reduced computation time
  - The number of basis dimensions was checked with the _gam.check_ function from _mgcv_
  - A smoother was not used for polygon size because the relationship was easily approximated with a log-linear relationship using a single slope term
  - All models were fit in _R_ using the _mgcv_ library, and figures were created with _ggplot2_ and _ggpubr_

- Most models still had a large amount of spatio-temporal autocorrelation in residuals
  - We have 50000 data points per field, so we ignored it

- To consider results from all field-level models while also accounting for field-level uncertainty, we used an approach similar to bootstrapping of hierarchical mixed effects models
  - The approach was as follows
  1. Extract single posterior sample (_rnorm_ in R) of smoother parameters from each field-level model using coefficient estimates and standard error
  2. Use posterior sample to create new simulated smoother from each field
  3. Fit model of all simulated smoothers, and save this meta-smoother
  4. Repeat 1000 times, and calculate coverage intervals (5-95% percentiles) on saved meta-smoothers
  - This gives credible intervals for the "average" smoother while accounting for field-level variability
